\section{Introduction}
\label{sec:intro}

Let $A$ be a commutative ring and $k\in\mathbb{Z}^+$.
Fix a \textbf{shape} $\vec{s}=(n_1,\ldots,n_k)\in(\mathbb{Z}^+)^k$ and set $n=\min(\vec{s})-1$.
We construct the diagonal simplicial tensor module $X_\bullet(\vec{s};A)$ (DSTM), a simplicial $A$-module
whose $p$-simplices are functions on cosimplicial index sets $I_p(\vec{s})\subseteq\mathbb{N}^k$.


This construction generalizes Quillen's diagonal on a double simplicial module~\cite{Quillen1966}
to $k$ commuting simplicial directions and is compatible with Dold--Kan normalization
in the sense of Dold, Kan, May, Weibel, and Goerss--Jardine~\cite{Dold1958,Kan1958,MaySimplicial,Weibel,GoerssJardine}.
In particular, the diagonal simplicial module $X_\bullet(\vec{s};A)$ is obtained from
$X_\bullet(\vec{s};\mathbb{Z})$ by base change, and the normalized Moore complexes satisfy
\[
N_\bullet\bigl(X_\bullet(\vec{s};A)\bigr)
  \cong N_\bullet\bigl(X_\bullet(\vec{s};\mathbb{Z})\bigr)\otimes_{\mathbb{Z}} A
\]
for all commutative rings $A$.


When $k=2$ and $s_1=s_2$, the normalized chain complex of $X_\bullet(\vec{s};A)$
recovers the combinatorial Laplacian of a finite graph, together with effective resistance
and Foster-type identities in the classical sense~\cite{Foster1949,KleinRandic1993,SpielmanSrivastava2011,Harris1992}.



We analyze the horn kernels
\[
R_{p,j}(X)=\bigcap_{i\neq j}\ker\bigl(d_i:X_p\to X_{p-1}\bigr),
\]
and show that their combinatorics controls the homotopy-theoretic behavior of the DSTM.
For the simplicial dimension $n=\min(\vec{s})-1$ we prove that $X_\bullet(\vec{s};A)$ is
a strict algebraic $n$-hypergroupoid (in the sense of Duskin--Glenn~\cite{Duskin1979,Glenn1982})
if and only if the tensor order satisfies $k=n$.
Thus the transition from non-unique to unique horn fillers occurs exactly at $p=n$,
and this threshold is determined by the tensor order.

The present work is not a general theory of Laplacians or effective resistance for arbitrary
hypergraphs or simplicial complexes.
Higher-order Laplacians and effective resistance for general simplicial complexes and simplicial networks
have been studied elsewhere~\cite{Chung1993,Eckmann1944,HorakJost2013,KookLee2018,GarciaRedondo2025};
our results are confined to the particular simplicial module $X_\bullet(\vec{s};A)$ and exploit its combinatorial rigidity.
We fix the functorial family of diagonal simplicial tensor modules
$X_\bullet(\vec{s};A)$ attached to a $k$-tensor, and work out the combinatorics, algebra,
geometry, and Hodge theory of this family in closed form.
All spectral statements in this paper are made for these DSTM complexes and for subcomplexes
generated by a single tensor.

In the remainder of this introduction we briefly outline the three main aspects of the paper:
combinatorics and classification of the DSTM, the algebraic geometry of generated submodules,
and combinatorial Hodge theory and effective resistance.

\subsection*{Combinatorics and Classification}

In the abelian setting of simplicial modules, horn fillers always exist (the Kan condition is automatic).
The obstruction to the uniqueness of fillers is captured by the horn kernel $R_{p,j}$.
Following terminology from Duskin--Glenn~\cite{Duskin1979,Glenn1982} and the $n$Lab,
a simplicial module is a \textbf{strict algebraic $n$-hypergroupoid} if fillers are unique
in dimensions strictly greater than $n$ ($R_{p,j}=0$ for $p>n$), and non-unique in dimension $n$ ($R_{n,j}\ne 0$).

We characterize $R_{n,j}$ combinatorially via a basis of ``missing indices.''
The Horn Non-Degeneracy Lemma proves $R_{n,j}\cap D_n=\{0\}$, where $D_n$ denotes the degenerate submodule in degree $n$,
yielding the decomposition $X_n = R_{n,j} \oplus D_n$.
Counting missing indices gives exact rank formulas for $R_{p,j}$ via inclusion--exclusion on the product sets
$I_p = \prod_a [M_a(p)]$.
In the constant shape case $n_a=n+1$ for all $a$, so $M_a(p)=p$ and $I_p=[p]^k$, and these rank formulas
specialize to finite differences of $x^k$ and can be expressed using Stirling numbers of the second kind
$\StirlingII{k}{m}$~\cite{StanleyEC1}.
By analyzing when the horn kernel vanishes, we obtain our main classification theorem:
$X_\bullet(\vec{s};A)$ is a strict algebraic $n$-hypergroupoid if and only if $k=n$.

\subsection*{Algebraic Geometry of Generated Submodules}

We work throughout with submodules generated by a single tensor $T\in X_n(\vec{s};K)$ over an infinite field $K$;
we do not attempt to parametrize arbitrary subcomplexes.
The generated submodule $\langle T \rangle$ is determined by its kernel sequence $(\ker f_{T,p})_p$.

We define a moduli map $\Psi$ from a Zariski-open locus $\mathcal{U} \subset X_n(\vec{s};K)$
to a product of Grassmannians.
The image $\mathcal{M}(\vec{s})$ lies within a closed incidence subvariety defined by simplicial compatibility conditions,
which we show are linear in the Segre--Pl\"ucker coordinates.
This reduces the classification of these generated submodules to the geometry of the index collision maps $\mathcal{I}_p$.

\subsection*{Combinatorial Hodge Theory}

Over a field of characteristic zero, we apply the standard combinatorial Hodge decomposition on a finite simplicial complex
(compare Eckmann and Horak--Jost~\cite{Eckmann1944,HorakJost2013}).
In the DSTM setting we combine this decomposition with the explicit rank formulas and contracting homotopy
to obtain a Foster-type identity in all degrees and to express Betti numbers and effective resistance in terms of the tensor data.

Since $X_\bullet$ is contractible, the Laplacian $\Delta_p$ is invertible for all $p\geq 0$.
Its inverse $G_p$ defines an effective resistance metric.
For generated submodules $\langle T \rangle$ with nontrivial homology, the projected Laplacian yields Betti numbers
and an effective resistance pseudometric on $\langle T\rangle$.

\subsection*{Organization}

Section~\ref{sec:dstm} establishes notation and the DSTM construction.
Section~\ref{sec:horns} characterizes missing indices and the horn kernel basis.
Section~\ref{sec:normalization} develops the Moore filler algorithm and normalization.
Section~\ref{sec:combinatorics} establishes rank formulas and the classification theorem.
Section~\ref{sec:hornnondegeneracylemma} proves the Horn Non-Degeneracy Lemma and the
decomposition $X_n = R_{n,j} \oplus D_n$.
Section~\ref{sec:dichotomy} develops the homology dichotomy and the hypergroupoid classification.
Section~\ref{sec:generated} develops the algebraic geometry of generated submodules.
Section~\ref{sec:hodge} develops Hodge theory and effective resistance for the DSTM
and its generated subcomplexes.
\ref{app:filtration} proves contractibility via an explicit equivariant homotopy.

